---
title: "Esame"
author: "Davide Puzzer"
date: "11/8/2021"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```
```{r}
install.packages("tidyverse")
install.packages("plotly")
```

#pacchetti
```{r}
library(dplyr)

library(tidyverse)

library(ggplot2)

library(plotly)

```

#import
```{r}
campionato_costruttori <- read.table("C:/Users/david/Desktop/Università/Scienza dei dati/f1db_csv/constructor_results.csv",header=TRUE, stringsAsFactors=TRUE, sep=",", na.strings="NA", dec=".", strip.white=TRUE)

constructors_standings <- read.table("C:/Users/david/Desktop/Università/Scienza dei dati/f1db_csv/constructor_standings.csv", header=TRUE, stringsAsFactors=TRUE, sep=",", na.strings="NA", dec=".", strip.white=TRUE)

costruttori <- read.table("C:/Users/david/Desktop/Università/Scienza dei dati/f1db_csv/constructors.csv", header=TRUE, stringsAsFactors=TRUE, sep=",", na.strings="NA", dec=".", strip.white=TRUE)

drivers_standings <- read.table("C:/Users/david/Desktop/Università/Scienza dei dati/f1db_csv/driver_standings.csv", header=TRUE, stringsAsFactors=TRUE, sep=",", na.strings="NA", dec=".", strip.white=TRUE)

drivers <- read.table("C:/Users/david/Desktop/Università/Scienza dei dati/f1db_csv/drivers.csv", header=TRUE, stringsAsFactors=TRUE, sep=",", na.strings="NA", dec=".", strip.white=TRUE)

lap_times <- read.table("C:/Users/david/Desktop/Università/Scienza dei dati/f1db_csv/lap_times.csv", header=TRUE, stringsAsFactors=TRUE, sep=",", na.strings="NA", dec=".", strip.white=TRUE)

pitstops <- read.table("C:/Users/david/Desktop/Università/Scienza dei dati/f1db_csv/pit_stops.csv", header=TRUE, stringsAsFactors=TRUE, sep=",", na.strings="NA", dec=".", strip.white=TRUE)

qualifiche <- read.table("C:/Users/david/Desktop/Università/Scienza dei dati/f1db_csv/qualifying.csv", header=TRUE, stringsAsFactors=TRUE, sep=",", na.strings="NA", dec=".", strip.white=TRUE)

races <- read.table("C:/Users/david/Desktop/Università/Scienza dei dati/f1db_csv/races.csv", header=TRUE, stringsAsFactors=TRUE, sep=",", na.strings="NA", dec=".", strip.white=TRUE)

risultati <- read.table("C:/Users/david/Desktop/Università/Scienza dei dati/f1db_csv/results.csv", header=TRUE, stringsAsFactors=TRUE, sep=",", na.strings="NA", dec=".", strip.white=TRUE)

stagioni <- read.table("C:/Users/david/Desktop/Università/Scienza dei dati/f1db_csv/seasons.csv", header=TRUE, stringsAsFactors=TRUE, sep=",", na.strings="NA", dec=".", strip.white=TRUE)

status <- read.table("C:/Users/david/Desktop/Università/Scienza dei dati/f1db_csv/status.csv", header=TRUE, stringsAsFactors=TRUE, sep=",", na.strings="NA", dec=".", strip.white=TRUE)


```
#area-test
```{r}
stagione_2018<-risultati %>%
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018) %>%
	select(driverRef,position,points)


```

#variabili

```{r}
campionato_costruttori
constructors_standings
costruttori
drivers_standings
drivers
lap_times
pitstops
qualifiche
races
risultati
stagioni
status

piloti
gare

australia
bahrain
shanghai
baku
montmelo
monaco
canada
francia
redbullring
silverstone
hockenheimring
hungaroring
spa
monza
singapore
sochi
suzuka
austin
messico
brasile
yasmarina

```

#piloti
```{r}
piloti<-risultati %>% 
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018) %>%
	slice(1:20) %>%
	select(driverId,forename,driverRef,nationality)

```

#calendario
```{r}
gare<-races %>%
	filter(year == 2018) %>%
	select(circuitId,round,name,date,raceId)

```

#classifica_gara_per_gara
```{r}
australia<- risultati %>%
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018, round == 1) %>%
	select(driverId,driverRef,position,points)

bahrain<- risultati %>%
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018, round == 2) %>%
	select(driverId,driverRef,position,points)

shanghai<- risultati %>%
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018, round == 3) %>%
	select(driverId,driverRef,position,points)

baku<- risultati %>%
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018, round == 4) %>%
	select(driverId,driverRef,position,points)

montmelo<- risultati %>%
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018, round == 5) %>%
	select(driverId,driverRef,position,points)

monaco<- risultati %>%
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018, round == 6) %>%
	select(driverId,driverRef,position,points)

canada<- risultati %>%
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018, round == 7) %>%
	select(driverId,driverRef,position,points)

francia<- risultati %>%
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018, round == 8) %>%
	select(driverId,driverRef,position,points)

redbullring<- risultati %>%
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018, round == 9) %>%
	select(driverId,driverRef,position,points)

silverstone<- risultati %>%
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018, round == 10) %>%
	select(driverId,driverRef,position,points)

hockenheimring<- risultati %>%
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018, round == 11) %>%
	select(driverId,driverRef,position,points)

hungaroring<- risultati %>%
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018, round == 12) %>%
	select(driverId,driverRef,position,points)

spa<- risultati %>%
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018, round == 13) %>%
	select(driverId,driverRef,position,points)

monza<- risultati %>%
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018, round == 14) %>%
	select(driverId,driverRef,position,points)

singapore<- risultati %>%
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018, round == 15) %>%
	select(driverId,driverRef,position,points)

sochi<- risultati %>%
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018, round == 16) %>%
	select(driverId,driverRef,position,points)

suzuka<- risultati %>%
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018, round == 17) %>%
	select(driverId,driverRef,position,points)

austin<- risultati %>%
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018, round == 18) %>%
	select(driverId,driverRef,position,points)

messico<- risultati %>%
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018, round == 19) %>%
	select(driverId,driverRef,position,points)

brasile<- risultati %>%
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018, round == 20) %>%
	select(driverId,driverRef,position,points)

yasmarina<- risultati %>%
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018, round == 21) %>%
	select(driverId,driverRef,position,points)
```

```{r}
lewis<- risultati %>%
  inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
  filter(driverId == 1 , year == 2018) %>%
  select(name,position,points,round)
lewis


```

#plot andamento di Lewis Hamilton
```{r}
ggplot(data=lewis,mapping=aes(x=round , y=reorder(position,desc(position)) , color=position)) +
  geom_point(mapping=aes(size=points))
 
```
```{r}
ggplot(data=lewis,mapping=aes(x=round , y=position , color=position)) +
  geom_point(mapping=aes(size=points)) +
  geom_line(size=1.15) +
  scale_y_continuous(trans = "reverse", breaks=unique(lewis$position))
```
#test lewis
```{r}
test_lewis<-ggplot(data=lewis,mapping=aes(reorder(name ,round), y=position)) +
  geom_point(size=6,color=lewis$position) +
  scale_y_continuous(trans = "reverse", breaks=unique(lewis$position)) +
  geom_text(data=lewis,label=posizioncine,color="white") 
  #geom_text()
  
print(test_lewis +
        ggtitle("Stagione 2018 Lewis Hamilton #44") +
        labs(y="Posizione al traguardo",x="Gare"))
```

#dati Sebastian Vettel
```{r}
sebastian<- risultati %>%
  inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
  filter(driverId == 20 , year == 2018) %>%
  select(name,position,points,round)
sebastian
```
#plotting sebastian vettel
```{r}
ggplot(data=sebastian,mapping=aes(x=round , y=position , color=position)) +
  geom_point(mapping=aes(size=points)) +
  geom_line(size=1.15) +
  scale_y_continuous(trans = "reverse", breaks=unique(sebastian$position))
```
```{r}
test_seby<-ggplot(data=sebastian,mapping=aes(reorder(name ,round), y=position)) +
  geom_point(size=6,color=sebastian$position) +
  scale_y_continuous(trans = "reverse", breaks=unique(sebastian$position)) +
  geom_text(data=sebastian,label=posizioncine_seby,color="white") 
  #geom_text()
  
print(test_seby +
        ggtitle("Stagione 2018 Sebastian Vettel #5") +
        labs(y="Posizione al traguardo",x="Gare"))
```
#tempi Hockenheimring 2018
```{r}
tempi_seb<-lap_times %>% 
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018,raceId == 999,driverRef == "vettel")
  #select(lap,time.x,position)

tempi_lewis<-lap_times %>% 
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018,raceId == 999,driverRef == "hamilton")
  #select(lap,time.x,position)

pitlewis<- pitstops %>%
    inner_join(drivers,by="driverId") %>%
    inner_join(races,by="raceId") %>%
  filter(year == 2018,raceId == 999,driverRef == "hamilton")

pitseby<- pitstops %>%
    inner_join(drivers,by="driverId") %>%
    inner_join(races,by="raceId") %>%
  filter(year == 2018,raceId == 999,driverRef == "vettel")
```


```{r}
#seby_time$time.x<-as.numeric(as.character(seby_time$time.x))

seby_time<-ggplot(tempi_seb,aes(lap,milliseconds)) + 
  geom_point() +
  geom_line(color="red")+
  scale_y_continuous(breaks = c(80000, 84000, 88000,92000), labels = c("1:20:000", "1:24:000","1:28:000","1:32:000"))
  
  
seby_time +
  ggtitle("Tempi sul giro Hockenheimring 2018 Sebastian Vettel #5") +
  xlab("Giro") + 
  ylab("Tempi sul giro")

fig<-ggplotly(seby_time)
fig
```
#tempi Lewis Hamilton
```{r}
lewis_time<-ggplot(tempi_lewis,aes(lap,milliseconds),color=position) + 
  geom_point() +
  geom_line(color="green") +
  scale_y_continuous(breaks = c(80000,100000,120000,140000), labels = c("1:20:000", "1:40:000","2:00:000","2:20:000"))
  #stat_summary(fun.y=sum,geom="line")
  
lewis_time +
  ggtitle("Tempi sul giro Hockenheimring 2018 Lewis Hamilton #44") +
  xlab("Giro") + 
  ylab("Tempi sul giro")

fig<-ggplotly(lewis_time)
fig
```

#area test
```{r}
ggplot(tempi_seb,aes(milliseconds,lap))+
  geom_line(stat="identity")
```
```{r}
t<-lap_times %>% 
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018,raceId == 999,driverRef == "vettel") %>%
  select(milliseconds)
l<-lap_times %>% 
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018,raceId == 999,driverRef == "vettel") %>%
  select(lap)

as.vector(t)
as.vector(l)

lol<-data.frame(l,t)

plot(lol,type="b",pch=19,col="red",xlab="Giri",ylab="Tempi")
#axis(side=1,at=t)
#axis(side=2,at=l)
box()
```

#campionato costruttori
```{r}

mercedes <- costruttori %>%
  filter(constructorRef == "mercedes")

ferrari <- costruttori %>%
  filter(constructorRef == "ferrari")

costr_2018_ferrari <- constructors_standings %>% 
	inner_join(costruttori,by="constructorId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018,constructorRef == "ferrari")

costr_2018_mercedes <- constructors_standings %>% 
	inner_join(costruttori,by="constructorId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018,constructorRef == "mercedes")

```

#grafico ferrari
```{r}
costruttoriferrari<-ggplot(costr_2018_ferrari,aes(reorder(name.y ,points), points))+
  geom_bar(stat="identity",fill="red")
costruttoriferrari
```
#grafico mercedes
```{r}
costruttorimercedes<-ggplot(costr_2018_mercedes,aes(reorder(name.y ,points), points))+
  geom_bar(stat="identity",fill="green")
costruttorimercedes
```
```{r}
bigs<- constructors_standings %>% 
	inner_join(costruttori,by="constructorId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018,constructorRef == c("ferrari","mercedes"))
```

```{r}
plotbigs<-ggplot(data=bigs)+
  geom_bar(
    aes(reorder(name.y ,points), points ,fill=Costruttori),
    position="dodge",
    stat="identity"
  )
plotbigs +
  ggtitle("Campionato costruttori Scuderia Ferrari - Mercedes AMG Petronas 2018") +
  xlab("Gare") + 
  ylab("Punti") +
  theme(plot.title = element_text(color = "blue", size = 12, face = "italic"))

  
  
plotbigs
```
#monza 2018
```{r}
tempi_kimi<-lap_times %>% 
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018,raceId == 1002,driverRef == "raikkonen")

tempi_ham<-lap_times %>% 
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018,raceId == 1002,driverRef == "hamilton")

tempi_misto<-lap_times %>% 
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018,raceId == 1002,driverRef == c("hamilton","raikkonen"))
```

```{r}
kimi_time<-ggplot(tempi_kimi,aes(lap,milliseconds)) + 
  geom_point() +
  geom_line(color="red")+
  scale_y_continuous(breaks = c(90000, 100000,110000,120000,130000,140000),
                     labels = c("1:30:000", "1:40:000","1:50:000","2:00:000","2:10:000","2:20:000"))
  
  
kimi_time +
  ggtitle("Tempi sul giro Monza 2018 Kimi Raikkonen #7") +
  xlab("Giro") + 
  ylab("Tempi sul giro")

#figura<-ggplotly(kimi_time)
#figura
```
```{r}
ham_time<-ggplot(tempi_ham,aes(lap,milliseconds)) + 
  geom_point() +
  geom_line(color="green")+
  scale_y_continuous(breaks = c(80000, 90000, 100000,110000,120000,130000,140000),
                     labels = c("1:20:00","1:30:000", "1:40:000","1:50:000","2:00:000","2:10:000","2:20:000"))
  
  
ham_time +
  ggtitle("Tempi sul giro Monza 2018 Lewis Hamilton #44") +
  xlab("Giro") + 
  ylab("Tempi sul giro")
```

#test
```{r}
mista<-ggplot(data=tempi_misto)+
  geom_line(
    aes(lap, milliseconds ,fill=number),
    color=tempi_misto$number,size=1.20,position="dodge",
    stat="identity")+
  scale_y_continuous(breaks = c(80000, 90000, 100000,110000,120000,130000,140000),
                     labels = c("1:20:00","1:30:000", "1:40:000","1:50:000","2:00:000","2:10:000","2:20:000"))+
  scale_colour_manual(name="Piloti",values=tempi_misto$number)
  
mista +
  ggtitle("Confronto Hamilton-Raikkonen Monza 2018") +
  xlab("Giri") + 
  ylab("Tempi") +
  theme(plot.title = element_text(color = "blue", size = 12, face = "italic"))
```
```{r}
monza<-ggplot(data=tempi_monza,mapping=aes(lap,milliseconds,color=Pilota))+
  geom_line(bindwidth=0.1,stat="identity",size=1.10)+
  scale_y_continuous(breaks = c(80000, 90000, 100000,110000,120000,130000,140000),
                     labels = c("1:20:00","1:30:000", "1:40:000","1:50:000","2:00:000","2:10:000","2:20:000"))+
geom_point()

  monza+
    scale_color_manual(values=c("aquamarine3","Red"))+
    ggtitle("Confronto Hamilton-Raikkonen Monza 2018") +
  xlab("Giri") + 
  ylab("Tempi") +
  theme(plot.title = element_text(color = "blue", size = 12, face = "italic"))
```
#Austin 2018

```{r}
tempi_austin_kimi<-lap_times %>% 
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018,raceId == 1006,driverRef == c("raikkonen"))
tempi_austin_lewis<-lap_times %>% 
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018,raceId == 1006,driverRef == c("hamilton"))
tempi_austin_max<-lap_times %>% 
	inner_join(drivers,by="driverId") %>%
	inner_join(races,by="raceId") %>%
	filter(year == 2018,raceId == 1006,driverRef == c("max_verstappen"))

tempi_austin <- rbind(tempi_austin_kimi,tempi_austin_lewis,tempi_austin_max)

```

```{r}
austin<-ggplot(data=tempi_austin,mapping=aes(lap,milliseconds,color=Pilota))+
  geom_line(bindwidth=0.1,stat="identity",size=1.10)+
  scale_y_continuous(breaks = c(100000,110000,120000,130000,140000),
                     labels = c("1:40:000","1:50:000","2:00:000","2:10:000","2:20:000"))+
  geom_point()

  austin+
    scale_color_manual(values=c("aquamarine3","Red","blue"))+
    ggtitle("Confronto Hamilton-Raikkonen-Verstappen Austin 2018") +
  xlab("Giri") + 
  ylab("Tempi") +
  theme(plot.title = element_text(color = "blue", size = 12, face = "italic"))
```
```{r}
pitkimilewis<- pitstops %>%
    inner_join(drivers,by="driverId") %>%
    inner_join(races,by="raceId") %>%
  filter(year == 2018,raceId == 1006,driverRef == c("raikkonen"))
```


#revisione hockenheim
```{r}
tempi_hock <- rbind(tempi_lewis,tempi_seb)
```

```{r}
hock<-ggplot(data=tempi_hock,mapping=aes(lap,milliseconds,color=Pilota))+
  geom_line(bindwidth=0.1,stat="identity",size=1.10)+
  scale_y_continuous(breaks = c(80000,100000,120000,140000),
                     labels = c("1:20:000","1:40:000","2:00:000","2:20:000"))+
  geom_point()

  hock+
    scale_color_manual(values=c("aquamarine3","Red"))+
    ggtitle("Confronto Sebastian Vettel - Lewis Hamilton Hockenheimring 2018") +
  xlab("Giri") + 
  ylab("Tempi") +
  theme(plot.title = element_text(color = "blue", size = 12, face = "italic"))
```
#revisione monza
```{r}
tempi_monza <- rbind(tempi_kimi,tempi_ham)
```

#tutti i pitsotops
```{r}
#hock
pitlewishock<- pitstops %>%
    inner_join(drivers,by="driverId") %>%
    inner_join(races,by="raceId") %>%
  filter(year == 2018,raceId == 999,driverRef == "hamilton")

pitsebyhock<- pitstops %>%
    inner_join(drivers,by="driverId") %>%
    inner_join(races,by="raceId") %>%
  filter(year == 2018,raceId == 999,driverRef == "vettel")

polechock<- qualifiche %>%
  inner_join(drivers,by="driverId") %>%
  inner_join(races,by="raceId") %>%
  filter(year == 2018,raceId == 999)

#monza

pitlewismonza<- pitstops %>%
    inner_join(drivers,by="driverId") %>%
    inner_join(races,by="raceId") %>%
  filter(year == 2018,raceId == 1002,driverRef == "hamilton")

pitkimimonza<- pitstops %>%
    inner_join(drivers,by="driverId") %>%
    inner_join(races,by="raceId") %>%
  filter(year == 2018,raceId == 1002,driverRef == "raikkonen")

polemonza<- qualifiche %>%
  inner_join(drivers,by="driverId") %>%
  inner_join(races,by="raceId") %>%
  filter(year == 2018,raceId == 1002)

#austin

pitlewisaustin<- pitstops %>%
    inner_join(drivers,by="driverId") %>%
    inner_join(races,by="raceId") %>%
  filter(year == 2018,raceId == 1006,driverRef == "hamilton")

pitkimiaustin<- pitstops %>%
    inner_join(drivers,by="driverId") %>%
    inner_join(races,by="raceId") %>%
  filter(year == 2018,raceId == 1006,driverRef == "raikkonen")

pitverstappenaustin<- pitstops %>%
    inner_join(drivers,by="driverId") %>%
    inner_join(races,by="raceId") %>%
  filter(year == 2018,raceId == 1006,driverRef == "max_verstappen")

poleaustin<- qualifiche %>%
  inner_join(drivers,by="driverId") %>%
  inner_join(races,by="raceId") %>%
  filter(year == 2018,raceId == 1006)



hock<- races %>%
  filter(year == 2018,raceId == 999 )
mo<-races %>%
  filter(year == 2018,raceId == 1002)
au<-races %>%
  filter(year == 2018,raceId == 1006)

pitlewishock
pitsebyhock

pitlewismonza
pitkimimonza

pitlewisaustin
pitkimiaustin
pitverstappenaustin

hock
mo
au
```

